% Submission: 4th eBPF Workshop (eBPF'26), co-located with ACM SOSP 2026
% Submission deadline: 2026-06-19
% Format: Research paper, 6 pages body + unlimited references
% Anonymity: double-blind for review submission

% \documentclass[sigconf,review,anonymous]{acmart}
\documentclass[sigplan,10pt,review,anonymous]{acmart}
\settopmatter{printfolios=true}
\pagestyle{plain}

% --- ACM metadata (anonymized for review; restore in camera-ready) ---
\setcopyright{none}
\acmConference[eBPF'26]{4th ACM SIGCOMM Workshop on eBPF and Kernel Extensions}{September 29, 2026}{Prague, Czech Republic}
\acmYear{2026}
\settopmatter{printacmref=false, printccs=false, printfolios=true}

\citestyle{acmnumeric}

% --- Packages ---
\usepackage{booktabs}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{pifont}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}

% --- Scoreboard marks for Table 4 (expressiveness) ---
% Use \partsym (not \partial) — \partial is the LaTeX kernel math symbol.
\newcommand{\yes}{\textcolor{green!55!black}{\ding{51}}}        % supported
\newcommand{\no}{\textcolor{red!70!black}{\ding{55}}}            % unsupported
\newcommand{\partsym}{\textcolor{orange!85!black}{$\sim$}}       % partial / awkward
\newcommand{\na}{\textcolor{gray}{\textemdash}}                  % not applicable

% --- Listings style for kunai DSL snippets ---
\lstdefinestyle{kunai}{
    basicstyle=\small\ttfamily,
    columns=fullflexible,
    breaklines=true,
    keepspaces=true,
}
% Tighten the vertical space above/below code listings.
\lstset{aboveskip=5pt, belowskip=3pt}

% --- Macros ---
\newcommand{\kunai}{Kunai\xspace}
\newcommand{\xdpninja}{xdp-ninja\xspace}
\newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}
\newcommand{\figref}[1]{Figure~\ref{#1}}

% Let column slack collect at the bottom instead of being stretched into
% the gaps around floats (acmart defaults to \flushbottom).
\raggedbottom

% Allow a little emergency stretch so long texttt identifiers do not
% poke past the column edge in the narrow two-column measure.
\setlength{\emergencystretch}{2em}

% Tighten the vertical space around floats and captions; acmart's
% defaults leave wide gaps between tables/figures and the body text.
\setlength{\textfloatsep}{10pt plus 2pt minus 2pt}
\setlength{\floatsep}{8pt plus 2pt minus 2pt}
\setlength{\intextsep}{10pt plus 2pt minus 2pt}
\setlength{\dbltextfloatsep}{10pt plus 2pt minus 2pt}
\setlength{\dblfloatsep}{8pt plus 2pt minus 2pt}
\setlength{\abovecaptionskip}{4pt}
\setlength{\belowcaptionskip}{2pt}

% Let a page hold more float area before deferring a float to a later
% page (the deferral is what leaves large gaps).
\renewcommand{\topfraction}{0.9}
\renewcommand{\bottomfraction}{0.8}
\renewcommand{\textfraction}{0.07}
\renewcommand{\floatpagefraction}{0.7}

\begin{document}

\title{Kunai: Toward a Verifier-Safe Layered DSL for Nested Encapsulation Packet Filtering on eBPF}

% --- Authors ---
% NOTE: the `anonymous' class option above suppresses all of the following
% and prints "Anonymous Author(s)" during review. These four blocks take
% effect only in the camera-ready (non-anonymous) build. Fill in real
% names/affiliations/emails before the final version.
\author{Takeru Hayasaka}
\affiliation{%
  \institution{Japan Advanced Institute of Science and Technology}
  \country{Japan}}
\additionalaffiliation{%
  \institution{BBSakura Networks Inc.}
  \country{Japan}}
\additionalaffiliation{%
  \institution{SAKURA internet Inc.}
  \country{Japan}}
\email{takeru.hayasaka@jaist.ac.jp}

\author{Satoshi Uda}
\affiliation{%
  \institution{Japan Advanced Institute of Science and Technology}
  \country{Japan}}
\email{zin@jaist.ac.jp}

\author{Ayako Hayasaka}
\affiliation{%
  \institution{Independent Researcher}}
\email{ternbusty.dev@gmail.com}

\author{Fourth Author}
\affiliation{%
  \institution{Kyoto University}
  \country{Japan}}
\email{kotani@media.kyoto-u.ac.jp}

\begin{abstract}
Packet capture and network debugging require selecting the packets
of interest early, inside the Linux kernel. However, pcap-filter,
tcpdump's filter language, has no syntax for nested
encapsulation, repeated occurrences of the same protocol, or the
variable-length lists and options whose field offsets change from
packet to packet. Running a filter in the kernel additionally
requires emitting bytecode that the Linux eBPF verifier accepts. For
such expressive filters this is difficult, since a naive lowering is
rejected. We present \kunai, a packet-filter DSL together with a compiler that
lowers its expressions to eBPF bytecode. A \kunai expression
describes a packet's layer structure and field conditions together,
and resolves each field's offset and type from
protocol definitions written in a subset of P4. The compiler records
every header's start offset at run time, bounds each variable-length
scan by a compile-time constant, and places a verifier-compliant
bounds check before every packet read. We evaluated \kunai on filters spanning GTP-U, SRv6, Geneve, and TCP
options. The generated bytecode was accepted by the verifier across
six kernels (Linux 6.1--7.0) on both the XDP and tc attach points, matched
and rejected the intended packets, and, where pcap-filter can express
the filter, ran within a few percent of pcap-filter in per-packet
processing cost. These
results show that \kunai enables the
Linux kernel to execute protocol-aware filters beyond what
pcap-filter's structured syntax can express.
\end{abstract}

\keywords{eBPF, XDP, packet filtering, domain-specific language, P4,
  kernel verifier, packet capture, network debugging}

\maketitle

\input{sections/01_introduction}
\input{sections/02_related}
\input{sections/03_design}
\input{sections/04_codegen}
\input{sections/05_evaluation}
\input{sections/06_limitations}
\input{sections/07_conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

\end{document}
